---
title: "R with public health data"
author: "Abhishek S D"
date: "11/05/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Basic R

1. To check which directory you are working in:

```{r}
getwd()
```

2. To import the data.
You need to change the “file” location to where you’ve stored the data set

```{r}
g <- read.csv(file = "D:/Courses/R/cancer data for MOOC 1.csv",header = TRUE, sep = ',') 
```

3. To have a look at the first few rows of our data set:

```{r}
head(g)
```

4. To inspect the `age` variable:

```{r}
g$age
```

5. To display a summary of the genders of our patients:

```{r}
table(g$gender)
```

6. To display a summary of the BMI of our patient:

```{r}
summary(g$bmi)
```

7. To display a summary of the smoking status of our patients:

```{r}
table(g$smoking) 
```

8. To display a summary of the exercise status of our patients:

```{r}
table(g$exercise)
```

9. To create a new variable `fruitveg`, which sums the daily consumption of fruit and veg of each patient:

```{r}
g$fruitveg <- g$fruit + g$veg 
```

10. To display a histogram of the ages of our patients:

```{r}
hist(g$age)
```

11. To create a new binary variable **five_a_day**, whether the patient eats at least 5 fruit or veg a day:

```{r}
g$five_a_day <- ifelse(g$fruitveg >= 5, 1, 0)
```

12. To summarise the **five_a_day** variable:

```{r}
table(g$five_a_day) 
```

13. To display a histogram of the daily fruit and veg consumption of our patients, including a title and proper axes:

```{r}
hist(g$fruitveg, xlab = "Portions of fruit and vegetables", main = "Daily consumption of fruit and vegetables combined", axes = F)

axis(side = 1, at = seq(0, 11, 1)) 

axis(side = 2, at = seq(0, 16, 2))
```

14. To create a new binary variable **healthy_BMI**, whether the patient has a healthy BMI or not:

```{r}
g$healthy_BMI <- ifelse(g$bmi > 18.5 & g$bmi < 25, 1, 0) 
```

15. To summarise **healthy_BMI**:

```{r}
table(g$healthy_BMI) 
```

16.  To run a chi-squared test to look for an association between eating five or more fruit and veg a day and cancer:

```{r}
chisq.test(x = g$five_a_day, y = g$cancer) 
```

16. To run a (two-tailed) t-test to see whether the mean BMI of those with cancer is different from the mean BMI of those without cancer: 

```{r}
t.test(g$bmi ~ g$cancer)
```

17. To run a (two-tailed) t-test to see whether the mean BMI of those with cancer is different from the mean BMI of those without cancer, where the variances are equal:

```{r}
t.test(g$bmi ~ g$cancer, var.equal = T)
```


18. To run a t-test to see whether the mean BMI of all patients is different from 25:

```{r}
t.test(g$bmi, mu = 25)
```

19. To run a chi-squared test to see whether there is an association between eating five or more fruit a day and having cancer:

```{r}
chisq.test(x = g$five_a_day, y = g$cancer)
```


20. To create a new binary variable, whether overweight or not according to their BMI:

```{r}
g$overweight <- ifelse(g$bmi >= 25, 1, 0)
```

21. To summarise the `overweight` variable:

```{r}
table(g$overweight) 
```

22. To run a chi-squared test to see whether there is an association between being overweight and cancer:

```{r}
chisq.test(x = g$overweight, y = g$cancer)
```

That's the end of basics in R.

